---
title: "Project4_Recommender_System"
author: "Chiranjeevi Konduru"
date: "2022-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(recommenderlab)
library(DT)
library(data.table)
library(reshape2)
```
```{r}
 myurl = "https://liangfgithub.github.io/MovieData/"
```

```{r}
# use colClasses = 'NULL' to skip columns
ratings = read.csv(paste0(myurl, 'ratings.dat?raw=true'), 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')
```


In movies.dat, some movie names contain single colon (:), so the method above does not work.

```{r}
movies = readLines(paste0(myurl, 'movies.dat?raw=true'))
movies = strsplit(movies, split = "::", fixed = TRUE, useBytes = TRUE)
movies = matrix(unlist(movies), ncol = 3, byrow = TRUE)
movies = data.frame(movies, stringsAsFactors = FALSE)
colnames(movies) = c('MovieID', 'Title', 'Genres')
movies$MovieID = as.integer(movies$MovieID)

# convert accented characters
movies$Title[73]
movies$Title = iconv(movies$Title, "latin1", "UTF-8")
movies$Title[73]

# extract year
movies$Year = as.numeric(unlist(
  lapply(movies$Title, function(x) substr(x, nchar(x)-4, nchar(x)-1))))
```

```{r}
users = read.csv(paste0(myurl, 'users.dat?raw=true'),
                 sep = ':', header = FALSE)
users = users[, -c(2,4,6,8)] # skip columns
colnames(users) = c('UserID', 'Gender', 'Age', 'Occupation', 'Zip-code')
```

```{r}
dim(users)
length(unique(ratings$UserID))
```

Unique Movies?

MovieID ranges from 1 to 3952, but not all of them are included in movies. Further, 177 movies are in movies but not rated in ratings.


```{r}
dim(movies)
```



```{r}
length(unique(ratings$MovieID))
```

```{r}
movies_not_rated = movies %>% 
  filter(!(MovieID %in% ratings$MovieID))
dim(movies_not_rated)
```


Dist of Ratings

Most of the ratings are in the 3-5 range.

```{r}
tmp = data.frame(Rating = 1:5, 
                 freq = as.vector(table(ratings$Rating)/nrow(ratings)))
ggplot(data = tmp, aes(x = Rating, y = freq)) +
  geom_bar(stat="identity", fill = 'steelblue', width = 0.6) + 
  geom_text(aes(label=round(freq, dig=2)), 
                vjust=1.6, color="white", size=3.5) +
  theme_minimal()
```


Ratings per User

Most users have left less than 500 ratings; about 9 users have had over 1300 ratings; one has 2314 ratings.

```{r}
tmp = ratings %>% 
  group_by(UserID) %>% 
  summarize(ratings_per_user = n()) 
summary(tmp$ratings_per_user)
stem(tmp$ratings_per_user)
sum(tmp$ratings_per_user > 500)
sort(tmp$ratings_per_user[tmp$ratings_per_user>1300])
```

```{r}
tmp %>%
  ggplot(aes(ratings_per_user)) +
  geom_bar(fill = "steelblue") + coord_cartesian(c(20, 500))
```


Combining users and tmp, you could further explore how ratings_per_user depends on Gender, Age, and Occupation of users.
```{r}
tmp = tmp %>% full_join(users, by = 'UserID')
```


Ratings per Movie

There are 31 movies that have received more than 2000 ratings. The most popular movie is “American Beauty (1999)”, followed by the “Star Wars” series. Throughout, popular means receiving many ratings; a popular movie may not be a highly-rated movie.

```{r}
tmp = ratings %>% 
  group_by(MovieID) %>% 
  summarize(ratings_per_movie = n(), ave_ratings = mean(Rating)) %>%
  inner_join(movies, by = 'MovieID')
summary(tmp$ratings_per_movie)
```


```{r}
tmp %>% 
  filter(ratings_per_movie > 2000) %>%
  arrange(desc = ratings_per_movie) %>%
  select(c("Title", "ratings_per_movie")) %>%
  print(n = 31)
```

```{r}
tmp %>% ggplot(aes(ratings_per_movie)) + 
  geom_bar(fill = "steelblue", width = 1) + coord_cartesian(c(1,1500))
```

The top ten highly-rated (based on their average ratings) among all movies that have received at least 1000 ratings.

```{r}
small_image_url = "https://liangfgithub.github.io/MovieImages/"
ratings %>% 
  group_by(MovieID) %>% 
  summarize(ratings_per_movie = n(), 
            ave_ratings = round(mean(Rating), dig=3)) %>%
  inner_join(movies, by = 'MovieID') %>%
  filter(ratings_per_movie > 1000) %>%
  top_n(10, ave_ratings) %>%
  mutate(Image = paste0('<img src="', 
                        small_image_url, 
                        MovieID, 
                        '.jpg?raw=true"></img>')) %>%
  select('Image', 'Title', 'ave_ratings') %>%
  arrange(desc(-ave_ratings)) %>%
  datatable(class = "nowrap hover row-border", 
            escape = FALSE, 
            options = list(dom = 't',
                          scrollX = TRUE, autoWidth = TRUE))
```

Dist of Genres

First cretae a bibary indicator for the 18 genres for each movie.


```{r}
genres = as.data.frame(movies$Genres, stringsAsFactors=FALSE)
tmp = as.data.frame(tstrsplit(genres[,1], '[|]',
                              type.convert=TRUE),
                    stringsAsFactors=FALSE)
genre_list = c("Action", "Adventure", "Animation", 
               "Children's", "Comedy", "Crime",
               "Documentary", "Drama", "Fantasy",
               "Film-Noir", "Horror", "Musical", 
               "Mystery", "Romance", "Sci-Fi", 
               "Thriller", "War", "Western")
m = length(genre_list)
genre_matrix = matrix(0, nrow(movies), length(genre_list))
for(i in 1:nrow(tmp)){
  genre_matrix[i,genre_list %in% tmp[i,]]=1
}
colnames(genre_matrix) = genre_list
remove("tmp", "genres")
```

Then we can output historograms of the 18 genres baesd on movies or based on ratings.

```{r}
data.frame(Genres = genre_list, 
                 Freq = as.vector(colMeans(genre_matrix))) %>% 
  ggplot(aes(reorder(Genres, Freq), Freq, fill = Freq)) + 
  geom_bar(stat = "identity") + 
    geom_text(aes(label = round(Freq, dig=2)), 
            position = position_stack(vjust = 0.5), 
            color="white", size=3) + 
  coord_flip() + 
  scale_colour_brewer(palette="Set1") + 
  labs(y = 'Frequency', x = 'Genre')
```


```{r}
tmp = ratings %>% 
  left_join(data.frame(MovieID = movies$MovieID, genre_matrix), 
            by = "MovieID") %>%
  select(-c("UserID", "MovieID", "Rating", "Timestamp"))
data.frame(Genres = genre_list, 
                 Popularity = as.vector(colMeans(tmp))) %>% 
  ggplot(aes(reorder(Genres, Popularity), Popularity, fill = Popularity)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = round(Popularity, dig=3)), 
            position = position_stack(vjust = 0.5), 
            color="white", size=3) + 
  coord_flip() + 
  labs(y = 'Popularity', x = 'Genre')
```

About half of the movies belong to only one genre; about one-third are categorized into two genres; a few are categorized into more than four genres. The movie “The Transformers” (1986) are categorized into six genres: Action, Animation, Children’s, Sci-Fi, Thriller, and War.

```{r}
tmp = rowSums(genre_matrix)
summary(tmp)

movies[which(tmp==6), ]
movies[which(tmp==5), ]
```


For illustration purpose only, let’s assume movies contains all the movies available to users from 1919 to 2000. Then we can compute the cumulative percentages of the 18 genres over year from 1919 to 2000 and store them in the 81-by-19 matrix tmp. For example, till 2000, users can access about 7.8% Action, 4.4% Adventure, 25% Drama, etc. A graphical display of such CDF over 10 generes are displayed below.


```{r}
# range(movies$Year) % 1919 to 2000
tmp = data.frame(Year = movies$Year, genre_matrix) %>%
  group_by(Year) %>%
  summarise_all(sum)
tmp[,-1] = apply(tmp[, -1], 2, cumsum)
tmp[,-1] = tmp[,-1]/sum(tmp[nrow(tmp), -1])
print(round(tmp[nrow(tmp),-1], dig=3))
```

```{r}
tmp = reshape2::melt(tmp, id.vars="Year") 
tmp %>%
  ggplot(aes(Year, value, group = variable)) +
  geom_area(aes(fill = variable)) + 
  geom_line(aes(group = variable), position = "stack")
```

